---
title: "4) EDA II"
---

This module involves a first exploratory analysis.

The idea is to have a first contact with the data.

```{r eval = FALSE}
# recommended packages
library(tidyverse)
library(lubridate)
library(viridis)
```


\


## What to Analyze?

Here are some EDA ideas.


### Tile Graph

Tile Graph of Year and Month

```{r eval = FALSE}
# without viridis color palette
crashes |> 
  count(ACCIDENT_YEAR, MONTH, name = "count") |> 
  ggplot(aes(x = ACCIDENT_YEAR, y = MONTH)) +
  geom_tile(aes(fill = count), linewidth = 0.5, color = "white") +
  scale_x_continuous(breaks = 2014:2023) +
  scale_y_continuous(breaks = 1:12)
```


```{r eval = FALSE}
# with viridis color palette
crashes |> 
  count(ACCIDENT_YEAR, MONTH, name = "count") |> 
  ggplot(aes(x = ACCIDENT_YEAR, y = MONTH)) +
  geom_tile(aes(fill = count), linewidth = 0.5, color = "white") +
  scale_fill_viridis() +
  scale_x_continuous(breaks = 2014:2023) +
  scale_y_continuous(breaks = 1:12)
```


### More Tile Graphs

- Month (x-axis) & Day of week (y-axis), for all years

- Month (x-axis) & Day of week (y-axis), for a given year

- Month (x-axis) & Day of week (y-axis), faceting by year

- Week (x-axis) & Day of week (y-axis)



### Types of Collision

```{r eval = FALSE}
# most common type of collision
crashes |> 
  count(TYPE_OF_COLLISION, name = "count", sort = TRUE)
```

```{r eval = FALSE}
# most common type of collision
crashes |> 
  count(TYPE_OF_COLLISION, name = "count", sort = TRUE) |> 
  ggplot(aes(y = reorder(TYPE_OF_COLLISION, count), x = count)) +
  geom_col() +
  labs(title = "Number of crashes (2014-2023)",
       y = "Type of crash")
```



### Type of collision per year

```{r eval = FALSE}
crashes |> 
  count(ACCIDENT_YEAR, TYPE_OF_COLLISION, name = "count") |> 
```

```{r eval = FALSE}
# normalized bar chart (Year, type of crash)
crashes |> 
  ggplot(aes(x = ACCIDENT_YEAR, fill = TYPE_OF_COLLISION)) +
  geom_bar(position = "fill")
```

```{r eval = FALSE}
# bar chart (Year, type of crash)
crashes |> 
  ggplot(aes(x = ACCIDENT_YEAR, fill = TYPE_OF_COLLISION)) +
  geom_bar()
```

```{r eval = FALSE}
# timeline (type of crash per year)
crashes |> 
  count(ACCIDENT_YEAR, TYPE_OF_COLLISION, name = "count") |> 
  ggplot(aes(x = ACCIDENT_YEAR, y = count, color = TYPE_OF_COLLISION)) +
  geom_line()
```


### Other ideas

- Most Common Type of Collision per Month

- Most Common Type of Collision per Hour

- Most Common Type of Collision per Hour and Day

- Most Common Type of Collision per County

- Most Common Type of Collision in a given City


